---
title: "Other Measures"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Other Measures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Redistricting is problem with many, many dimensions. This vignette introduces some useful measures related to redistricting, but with smaller categories. See the vignette "Using `redistmetrics`" for the bare-bones of the package. 

We first load the `redistmetrics` package and data from New Hampshire. For any function, the `shp` argument can be swapped out for your data, `rvote` and `dvote` for any two party votes, `black_pop` for any group population, `pop` for the total population, and the `plans` argument can be swapped out for your redistricting plans (be it a single plan, a matrix of plans, or a `redist_plans` object).  

```{r setup}
library(redistmetrics)
data(nh)
```


## Competitiveness

### Talismanic Compactness

This is a measure which offers a balance between competitiveness across the state and competitiveness within inidividual districts.

Talismanic Compactness can be computed with

```{r}
compet_talisman(plans = nh$r_2020, shp = nh, rvote = nrv, dvote = ndv)
```

where `nrv` and `ndv` are averages of votes. (In general, you want to compute these scores over many elections and average them.)

## Segregation

### Dissimilarity

Dissimilarity describes how similar the demographic proportions in districts are to the total state population's demographics.

Dissimilarity can be computed with:

```{r}
seg_dissim(plans = nh$r_2020, shp = nh, group_pop = black_pop, total_pop = pop)
```

## Incumbents

### Incumbent Pairings

We compute incumbent pairings as the number of incumbents who are placed into a district with other incumbents beyond those allowed. Formally, this is:

$$\textrm{Inc. Pairs} = \sum_{d = 1}^{\textrm{ndists}}\max(0, n^{(d)}_{inc} - 1)$$

We do not have incumbent data included for New Hampshire. As such, we create fake incumbent data.

```{r}
fake_inc <- rep(FALSE, nrow(nh))
fake_inc[c(1, 2)] <- TRUE
```

This would indicate that there are only incumbents in the first two rows of the data.

Incumbent pairings can be computed with:

```{r}
inc_pairs(plans = nh$r_2020, shp = nh, inc = fake_inc)
```


